Hierarchical storage management in user space for neuroimaging applications

### 2024-03-14 paper/sea-neuro/paper-neuro.tex

For both the pipelines and the datasets, we noticed that obtained speedups
decreased with increased parallelism. We did not modify the default parallelism
level of the applications in our experiments. As a result, each application
process was likely attempting to utilize all cores, resulting in longer wait times
with the increase in parallelism. With this unintentional additional wait time that
is added to the applications due to compute contention, the benefits of using
Sea becomes more limited.

@@

Baseline performance without busy writers was comparable to that of Sea's (p=0.7, two-sample unpaired t-test) as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
busy writers, the makespan obtained with Sea was smaller than baseline (p<1e-04, two-sample t-test). We did occasionally observe slowdowns from using Sea. These slowdowns
may arise from the initial read of the data, as they appeared to occur less
frequently with the SPM pipeline, or due to increased CPU contention caused by Sea's rapid I/O.

@@

    While on the production cluster, Sea and Baseline generally performed similarly,
    it is impossible for cluster users to know if Lustre's
    performance will be deteriorated during the lifetime of their application
    execution, nor to predict how such a deterioration would impact the
    application. The fact that Sea and Baseline perform similarly in the average
    case is an indication that runtime is generally not deteriorated by Sea, as
    also demonstrated by comparing Sea's performance with tmpfs. Therefore, Sea
    is safe to use in the event where Lustre performance is sufficiently
    degraded.

    Although Sea usage did result in performance degradation at times, the magnitude
    of the degradation was less than that of possible speedups
    obtained. Our controlled experiments demonstrated that using Sea with a
    degraded Lustre file system can result in average speedups of up to
    2.5$\times$. Such results indicate that the risk of a slowdown with Sea are
    less important that the potential speedups that can arise.

@@

    Our results expose the complexities of performance predictions on HPC clusters. We presented results on a dedicated cluster
    where we had
    control over the global cluster usage and were able to demonstrate that
    significant speedups could be obtained when all the Lustre storage devices were
    busy writing other data concurrently. However, even these experiments were
    limited in demonstrating potential speedups with Sea as there were many
    controlled variables, some of which may vary from cluster to cluster. For
    example, the production cluster had a total of 977 nodes and only 38 disks associated with
    their Lustre scratch file system. We can imagine that if 900 of the nodes
    are busy writing to the scratch file system, we would observe a speedup
    larger than what has been reported on our dedicated cluster with only 8 compute
    nodes and 44 disks associated to Lustre. However, these are not the only
    factors that can affect reported results, others can include when the
    experiment were executed, what kind of applications were being executed
    alongside the experiment, scratch file system usage during experiments, and
    what were the application read and write patterns.

>>>>

For both the pipelines and the datasets, we noticed that obtained speedups
decreased with increased parallelism. We did not modify the default parallelism
level of the applications in our experiments. As a result, each application
process was likely attempting to utilize all cores, resulting in longer wait times
with the increase in parallelism. With the added compute contention overhead, the benefits of using
Sea becomes more limited.

@@

Baseline performance without busy writers was comparable to that of Sea's (p=0.7, two-sample unpaired t-test) as can be seen in Figure~\ref{fig:seaneuro:slashbin}. However, with
busy writers, the makespan obtained with Sea was smaller than baseline (p<1e-04, two-sample t-test). Occassional slowdowns did occur with the use of Sea. These slowdowns
may arise from the initial read of the data, as they appeared to occur less
frequently with the SPM pipeline, or due to increased CPU contention caused by Sea's rapid I/O.

@@

    It is unrealistic for cluster users to know whether Lustre's
    performance will be deteriorated during the lifetime of their application
    execution, nor to predict how such a deterioration would impact the
    application. The fact that Sea and Baseline perform similarly in the average
    case is an indication that runtime is generally not deteriorated by Sea, as
    also demonstrated by comparing Sea's performance with tmpfs. Therefore, Sea
    is safe to use in the event where Lustre performance is sufficiently
    degraded.

    Although Sea usage did result in performance degradation at times, the magnitude
    of the degradation was less than that of possible speedups
    obtained. Our controlled experiments demonstrated that using Sea with a
    degraded Lustre file system can result in average speedups of up to
    2.5$\times$, which is likely to benefit scientific application deployments in production.

@@

    Our results expose the complexities of performance predictions on HPC clusters. We presented results on a dedicated cluster
    where we had
    control over the global cluster usage and were able to demonstrate that
    significant speedups could be obtained when all the Lustre storage devices were
    busy writing other data concurrently. However, even these experiments were
    limited in demonstrating potential speedups with Sea as there were many
    controlled variables, some of which may vary from cluster to cluster. For
    example, the production cluster had a total of 977 nodes and only 38 disks associated with
    their Lustre scratch file system. We can imagine that if 900 of the nodes
    are busy writing to the scratch file system, we would observe a speedup
    larger than what has been reported on our dedicated cluster with only 8 compute
    nodes and 44 disks associated to Lustre. However, these are not the only
    factors that can affect reported results, others can include when the
    experiment was executed, what kind of applications were being executed
    alongside the experiment, scratch file system usage during experiments, and
    what were the application read and write patterns.
